Collaborative Visual Analysis on the Web with RCloud

### 2015-06-16 relatedwork.tex
yes

Previous work has identified some requirements for supporting the
visual analytics process, including the work of multiple users.

\paragraph*{Social Data exploration and Analysis}
ManyEyes~\cite{Viegas:2007:MAS} was a landmark system for the
crowdsourced creation and publication of data
visualizations. Although ManyEyes supported a fixed set of
visual encodings, the system's success was a precursor to more
sophisticated solutions and an early indication that the world-wide
web was a suitable platform for data visualization. One of the main
challenges we faced in designing RCloud was providing an experience
for \emph{consuming} visualizations as seamless as ManyEyes's, while
sacrificing as little as possible on the generality of the analyses
themselves.

\paragraph*{Notebooks as a medium for data analysis dissemination}
The concept of a ``notebook'' as we use it here can be traced all the
way back to Knuth's literate programming~\cite{Knuth:1984:LP}. In
literate programming, a comprehensive description in prose of the
behavior of the program is ``weaved'' together with the source code,
yielding both an executable program and a human-readable document.
A notebook represented as a collection of short, executable cells,
originated with Mathematica, and in R, literate programming is
supported by packages such as knitr and RMarkdown~\cite{Xie:2013:DDW}.
Project Jupyter~\cite{jupyter} (originally implemented as part of
IPython) offer some notebook features, but lack a transparent
mechanism for sharing and deployment in multiple-user settings.
%
Further afield, {Electronic Lab Notebooks} are applications for organizing
and sharing data from scientific lab experiments\cite{Rubacha:2011:ELN}.
In a sense, we hope to adapt and extend this concept to the work of
visual analytics teams.
%
Although RStudio offers publication of literate R programs as a free
service on their website, the workflow is somewhat disconnected from
the development of those programs. Once they're uploaded, it's hard
for other users to build off of the work published (or even for the
original author to update new versions). In other words, RStudio
handles \emph{publication}, but not \emph{deployment} and sharing.

>>>>

Previous work has identified key requirements for supporting the
visual analytics process, including by multiple users.

\paragraph*{Social Data exploration and Analysis}
ManyEyes~\cite{Viegas:2007:MAS} was a landmark system for the
crowdsourced creation and publication of data visualizations.
Although ManyEyes supported only a fixed set of visual encodings,
the system's success was an early indication of the potential to
combine social media on the web with visual analytics.

\paragraph*{Notebooks as a medium for data analysis dissemination}
The concept of a ``notebook'' as we use it can be traced
back to Knuth's literate programming~\cite{Knuth:1984:LP}. In
literate programming, a prose description of the behavior of
a program is ``weaved'' with its source code, yielding both an
executable program and a human-readable document.
A notebook represented as a collection of small executable cells,
originated with Mathematica, and in R, literate programming is
supported by packages such as knitr and RMarkdown~\cite{Xie:2013:DDW}.
Project Jupyter~\cite{jupyter} (originally implemented as part of
IPython) offers some notebook features, though it lacks a transparent
mechanism for multi-user sharing and deployment.
%
Further afield, {Electronic Lab Notebooks} are applications for organizing
and sharing data from scientific lab experiments\cite{Rubacha:2011:ELN}.
In a sense, we hope to adapt and extend this concept to the work of
visual analytics teams.
%
\stephen{RStudio is highly acclaimed for its...}
Although RStudio offers publication of literate R programs as a free
service on their website, the workflow is somewhat disconnected from
the development of those programs. Once they're uploaded, it's hard
for other users to build off of the work published (or even for the
original author to update new versions). In other words, RStudio
handles \emph{publication}, but not \emph{deployment} and sharing.
